---
title: "Reproducible Research | Course Project 1"
author: "Olga Nevinchana"
date: "3/18/2020"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Opening libraries

```{r}
  
  library(dplyr)
  library (ggplot2)
  library(knitr)
  
```
  
# Code for reading in the dataset and/or processing the data

```{r}

  setwd("~/Documents/Data Science Specialisation/Rerpoducible_Research")
  activity <- read.csv("./activity.csv")

```
 
```{r}

activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
      
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

```
  
```{r}

activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
  
activity$daytype <- sapply(activity$date, function(x) {
  if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
  {y <- "Weekend"} else 
  {y <- "Weekday"}
  y
  })

```
  
```{r}

summary(activity)

```
  
# Mean total number of steps taken per day

## Total number of steps taken per day.
  
```{r}

    by_date <- aggregate(steps~date, activity, sum, na.rm = TRUE)

```
  
## Histogram of the total number of steps taken each day.

```{r}

ggplot(by_date, aes(x = steps)) +
      geom_histogram(col = "white", fill = "orange", breaks = seq(0,22500, by=2500)) +
      ggtitle("Total number of steps taken per day") + 
      xlab("Steps") +
      ylab("Number of days")

```

## Mean and median of the total number of steps taken per day

```{r}

mean(by_date$steps, na.rm = TRUE)
median(by_date$steps, na.rm = TRUE)
    
```

# Average daily activity pattern

## Time series plot of the average number of steps taken

```{r}

    by_interval <- aggregate(steps~interval, activity, mean, na.rm = TRUE)
    
    ggplot(by_interval, aes(x = interval, y = steps)) +
      geom_line(col = "darkgreen") +
      ggtitle("Average daily activity pattern") +
      ylab("Average number of steps")
    
```

## 5-minute interval that contains the maximum number of steps

```{r}

    time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
    max_interval <- which.max(time_series)
    names(max_interval)
    
```

# Imputting missing values

## Total number of missing values in the dataset

```{r}

  summary(activity)

```

## Strategy for filling in all of the missing values in the dataset.

```{r}

    activity_nomissing <- activity
    for (i in unique(activity_nomissing$interval)) {
      activity_nomissing$steps[is.na(activity_nomissing$steps) & activity_nomissing$interval == i] <- round(mean(activity_nomissing$steps[activity$interval == i], na.rm = T))
    }

```

## Histogram of the total number of steps taken each day & mean and median total number 
## of steps taken per day.

```{r}

    by_date_nomossing <- with(activity_nomissing, aggregate(steps, by = list(date), FUN=sum, na.rm = TRUE))
    names(by_date_nomossing) <- c("Date", "Steps")
    
```

```{r}

    ggplot(by_date_nomossing, aes(x = Steps)) +
      geom_histogram(col = "white", fill = "purple", breaks = seq(0,22500, by=2500)) +
      ggtitle("Total number of steps taken per day") + 
      xlab("Steps") +
      ylab("Number of days")

```

```{r}

  mean(by_date_nomossing$Steps, na.rm = TRUE)
  median(by_date_nomossing$Steps, na.rm = TRUE) 
    
```

# Differences in activity patterns between weekdays and weekends.

```{r}

  by_interval_daytype <- aggregate(steps~interval + daytype, activity, mean, na.rm = TRUE)
   
     ggplot(by_interval_daytype, aes(x = interval , y = steps, color = daytype)) +
      geom_line() +
      labs(title = "Average daily steps by wekkday", x = "Interval", y = "Average number of steps") +
      facet_wrap(~daytype, nrow = 2, ncol = 1)
    
```
